---
title: "Exploratory Data Analysis: An objective analysis of Donald Trump's facebook activity"
author: "James Dayhuff"
date: "October 22, 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Data origins

This data was collected by another individual using open source technology such as python and git. The data was then published to the public for free analysis.
The data can be found at this location:
https://docs.google.com/spreadsheets/d/1mLO7SFqHmUaZEpp87cwkM0luJutSwmwKMx7kaM9348U/edit

## Cleaning the data

The data came in an xlsx format via a public google sheet. Once the data was downloaded to my local file system, I converted it to a csv file and attempted to read it into R-Studio. I immediately ran into problems with the formatting of the data. The biggest issue was the "status_message" column, which included facebook status messages with all the punctuation included. Read.csv() misinterpreted the punctuation in the status messages and it threw off the formatting.
After removing the punctuation in excel with the "find and replace" functionality, I easily imported the data into R-Studio.

## About the data
High level overview of the data

```{r overview}
tfb = read.csv("trump_fb_analysis.csv")
summary(tfb)
```

## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
